Regression Cube Analysis of Cohort Study Data

### 2015-03-31 draft.tex
added Steffis revision

\textbf{Data Security.}
%%
Security issues are raised by uploading data into an online service such as our prototype.
%%
The use of epidemiological data is preceded by a detailed description of the analysis purpose and has to be approved by ethics committees.
%%
Preventive steps have to be taken to restrict access to unauthorized subjects.
%%
%We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
%%
We apply a SHA-256 hashing on the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
%%
Data sets are deleted from the server after closing a session.

@@

\section{Application} \label{application}
%%
\begin{figure*}[htb]
 \centering
 \includegraphics[width=7.0in]{figures/application}
 \caption{
 The analysis of the numerical and dichotomized target feature depicting liver-fat values yields similar results (left).
 %%
 In (a), hot spots for somatometrical features with high correlations were found.
 %%
 High correlations were also found for features depicting \emph{hepatic steatosis} (b).
 %%
 A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D mosaic plot.
 %%
 The hypothesis-free analysis of the breast fat data set (right) towards the \emph{parenchyma tissue percentage} of the breast displays correlations between \emph{age}, \emph{body fat}, \emph{hip circumference} as well as \emph{menstrual period}.
 }
  \label{fig:Application}
\end{figure*}
%%
In this section, we describe how we applied the \emph{Regression Cube} to two epidemiological data sets.
%%
The hepatic steatosis data set was analyzed using data mining algorithms, yielding risk groups, which we now analyze further.
%%
Also, we try to reproduce the prior results from this analysis as proof-of-concept of our method.
%%
The breast fat data set is the foundation for an explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
%%

@@

% OK so?
The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Hepatic Steatosis Data Set}
We employ the data set used in \cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
%%
The dichotomous target feature is derived from the liver fat concentration measured using MRI scans.
%%
% ok?  to indicate absence or presence of the disease.
Liver fat concentrations of no more than 10\% are mapped to the 'negative' class; values greater than 10\% are mapped to the 'positive' class to indicate absence or presence of the disease.
%%
The data set contains labels for 578 participants for which the MRI data was available to that point.
%%
The MRI scans for each subject have been introduced in SHIP-2, so the target feature is only present for one of three study waves.

@@

In \cite{Niemann2014}, the authors show different class distributions of liver fat concentrations of women and men.
%%
For women an association between age and liver fat was identified.
%Also, for women they identify that there is a association between age and liver fat concentration.
%%
An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
%%
% ist das richtig? nicht female<52,male, female>52?
Based on these observations, we perform our analysis on three sub-populations: \emph{females}, \emph{males} and \emph{females} older than 52 years.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between parenchyma tissue proportion in the female breast towards other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

The data comprises 1,186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
%%
It contains 231 dimensions, holding information about somatometric features (e.g., body size or weight) consumption behavior (e.g., alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women specific features (e.g., number of born children or contraception type) as well as mammography features (e.g., fat content or parenchyma tissue proportion to volume).
% It contains 231 dimensions, holding information about
% \begin{itemize}
% 	\item \emph{somatometric features}, e.g., body size, weight or BMI,
% 	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume,
% 	\item \emph{personal history}, e.g., occupation, marital status,
% 	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
% 	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
% 	\item \emph{mammography features} derived from the image data.
% \end{itemize}
%%
The latter were derived from manual segmented MRI data for each subject carried out by radiologists.
%%
%%
%\com{Describe Parenchym tissue and its relation to breast cancer.}
%\com{Describe how the data was acquired and preprocessed}

@@

All features related to the mammography attributes were stored in a additional file.

@@

We refer to each analysis step regarding to their belonging in the VA-Mantra (recall Sec.~\ref{sec:Workflow}).

@@

Therefore, \emph{UN} started the [\textbf{1.AF}] step using the dichotomized MRI fat liver concentration using the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.

@@

The Zoom, Filter and Analyze Further Step [\textbf{3.ZF}] was realized by slicing through the cube using the mouse input to inspect the hot spots.

@@

Analyzing the mosaic plot [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator-features as well as laboratory values, such as \emph{creatinine} (used renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.

@@

Due to the difference between our regression-model approach and the decision-tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features are not expected.

@@

\textbf{Analysis of non-discretized target feature.}
%%
%Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
Since our method can assess numerical target features, the analysis was conducted again for the not-dichotomized target using the same formula.
%%
The 3D cube representation already showed lower $R^2$ values in general. However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
%%
The correlation hot spots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
%%
We assume that the bias introduced by dichotomizing the fat liver content enforces the findings towards liver diseases, while using the numerical features is less expressive.

\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hot spot was always observable in the [\textbf{2.SI}] and [\textbf{3.ZF}] incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
The literature described relations between \emph{IL-6} and liver cancer \cite{He2013} as well as chronic liver diseases \cite{Streetz2003}.
%%
For mice, strong effects of \emph{IL-6} with hepatic steatosis were described \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
The analysis aims to find relationships on the breast fat data using mammography analysis features.
%%
Relationships between the share parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%
The [\textbf{1.AF}] was started by \emph{KH} using the default formula for hypothesis-free analysis ($Z \sim X + Y$).
%%
At first, she was interested in correlations with the \emph{parenchyma tissue} percentage, which was selected through the drop-down for the z-axis [\textbf{2.SI}].
%%
She observed strong correlations with \emph{age}, \emph{body fat percentage}, \emph{hip} and \emph{waist circumference} as well as \emph{menstrual period} or \emph{pregnancy status} as expected (Fig.~\ref{fig:Application} right).
%%
Women with higher \emph{body fat} also have a larger \emph{breast fat percentage}, which also correlates with other somatometric features.
%%
\emph{Age} is a strong influencing factor, as breast tissue and subsequently the parenchyma tissue degrades over time.
%%
\emph{KH} proceeded using [\textbf{3.ZF}] and [\textbf{4.DD}] to check for relationships for different target features, such as current \emph{hormone replacement therapy}, \emph{BI-RADS} (classification of the mammography findings) as well as different diseases, such as \emph{diabetes} or \emph{gout}.
%%
She observed relationships matching her expectations and expert knowledge.
%%
One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} towards \emph{spiral contraception} ($R^2$ of $0.77$).
%%
\emph{KH} proceeded with a detailed analysis of the parenchyma tissue.
%\emph{KH} derived at the hypotheses: \emph{Somatometric features impose the major influences on the parenchyma tissue.}
%%

\textbf{Detailed Breast Parenchyma Analysis.}
%%
The analysis was conducted by calculating the cube $Parenchyma\_Percentage \sim X + Y + Z$ [\textbf{1.AF}].
%%
Using the 3D cube, \emph{KH} observed several hotspots [\textbf{2.SI}].
%%
Navigating to them using the slicing facility of the 3D visualization [\textbf{3.ZF}] highlighted features of high influence, such as image-derived features, as \emph{glandular tissue density} and \emph{parenchyma segmentation} metrics.
%%
Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. to its strong influence to the parenchyma tissue.
%%
A surprising find were the strong correlations with \emph{kidney disorder} ($R^2$ values around $0.9$).
%%
The [\textbf{4.DD}] analysis however showed only 8 subjects with \emph{kidney disorder}.
%%
Too few subjects impose the risk of a biased finding.
%%
The correlation was noted and will be investigated further in the near future using a extensive data set.
%%
Lastly \emph{KH} assessed the influence of contraception related features, such as use \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
%%
Other consumption behavior features, such as \emph{alcohol intake} also yield no elevated $R^2$ values.
%%
\emph{KH} remarked that these features are believed to have an impact on the parenchyma tissue, but they are less reliability since they are self-reported.
%%
%mamma anamnesis breast fed, contraception (spirale, pille), alcohol intake
%Lastly, \emph{KH} assessed the influence of image-derived features on the parenchyma tissue.
%%
%She observed expected results 
%Using the slicing mode to fast 
%Hypotheses
%Detailed analysis: Cholesterin, influence of image-derived feautres, classification of the BI-Rads (how well did the manual segmentation perform)
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
Our method was well received among the domain experts.
%%
For the first time, they were able to derive an overview visualization custom tailored to underlying assumptions.
%%
\emph{KH} noted the ease of use, which "\emph{converts data sets into a feasible form}".
%%
She highlighted the efficiency of combining fast target feature selection with visually highlighting interesting results, enabling rapid analysis cycles.
%%
To get nearly similar results she would had to spend hours using \texttt{SPSS} and potentially missed interesting hot spots during this process.
%She noted that she usually would have to spend hours using \texttt{SPSS} to assess the data in the detail we presented.
%%
\emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
%%
%TI commented on the

@@

\textbf{Overview Visualizations are Preferred over Black-box Methods.}
%%
Explorative analysis based on the data gains importance in epidemiology with increasing data set complexity.
%While the classical epidemiological workflow represents the foundation of the research in this area, explorative analysis based on the data gains importance.
Results from automatic `black-box' methods, such as data mining algorithm are more often obscure to the expert.
%%
Findings and hypotheses derived through overview visualizations, however, are met with more confidence, because the users actually observed the behavior themselves.
%%
The participation and steering of the analysis using human pattern-detection and expert knowledge is preferred.
%%
%Incorporating data mining algorithms with overview visualizations of statistical measures allow for reasoning about risk factors using underlying epidemiological data sets rather than predefined hypotheses.
%%
%We  a higher confidence of an expert in a result he or she actually observed in the data rather 
%One major advantage we observed is the confidence of an expert in the results he or she actually observe.
%%
Observing \emph{expected} correlations matching the expert knowledge strengthens the confidence in the method and, subsequently, in the hypotheses generated from unanticipated relationships.
%Observing \emph{expected} correlations, which match the expert knowledge, strengthens the confidence in the method and, subsequently, in the hypothesis generated from unanticipated relationships.
%%
%\com{Include the human into the analysis provides more confidence into the result than data mining algorithms. People believe more in what they see, rather than in what they are presented by an algorithm. Regression cubes incorporate this by showing known associations as well as new ones. Preattentive feature display (either through fast slicing through a cube or by showing hot spots) helps.}

%%
\textbf{Using non-discretized features reduces information bias.}
%%
Discretization reduces the information space and introduces bias into the data and is therefore avoided in epidemiological research when possible.
%%
In contrast to many data mining algorithms, our method allows to use the concurrent analysis of heterogeneous data types.
%%
Investigations of the the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results but slightly differing when it comes to details. %, but differences  %differing in detail.
%%
The overall explanatory power towards the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the results towards the expected result.
%%
%Our investigation regarding the hepatic steatosis data set confirms differences in results
%%
%\com{Comment on the lesson learned from the last VAST Paper - "Usage of different categorizations depending on expected outcome." Now we have continuos data.}
%%

\textbf{Attention steering is crucial.}
%%
Important events have to be highlighted in overview visualizations to steer the users attention to interesting parts of the data.
%%
Poor guidance potentially leads to overlooked relationships.
%%
We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
%%
%Foe example, formulas with a static target feature (e.g. $Breast\_Cancer ~ X + Y + Z$) are rendered by highlighting differences rather than absolute values in the 3D cube visualizations (see Fig.~\ref{fig:Application} left) by comparing local values with the mean along the z-ray.
%%

@@

\section{Summary and Conclusion}
%%
We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Model expert knowledge using regression formulas allow for \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D mosaic plot for each cube slice.
%%
Details on demand for each model allow for detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.

@@

We published all associated code and provide an analysis platform open to heterogenous data types.
%%
We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.

>>>>

\textbf{Data Security.}
%%
Security issues are raised by uploading data into an online service such as our prototype.
%%
The use of epidemiological data is preceded by a detailed description of the analysis purpose and has to be approved by ethics committees.
%%
Preventive steps have to be taken to restrict access to unauthorized subjects.
%%
%We decided against a user account system because it reduces the ease of use largely without having a direct advantage for the user.
%%
We calculate a SHA-256 hash to derive the data set name using the data contents and disable directory listings on the web server to avoid data set downloads.
%%
Data sets are deleted from the server after closing a session.

@@

\section{Application} \label{application}
%%
\begin{figure*}[htb]
 \centering
 \includegraphics[width=7.0in]{figures/application}
 \caption{
 The analysis of the numerical and dichotomized target feature depicting liver fat values yields similar results (left).
 %%
 In (a), hotspots for somatometrical features with high correlations were found.
 %%
 High correlations were also found for features depicting \emph{hepatic steatosis} (b).
 %%
 A high correlation between \emph{Interleukin-6}, \emph{hepatic steatosis}, \emph{GGT} and \emph{Lipase} (highlighted using arrows) was revealed during the analysis using the 2D mosaic plot.
 %%
 The hypothesis-free analysis of the breast fat data set (right) w.r.t. the \emph{parenchyma tissue percentage} of the breast displays correlations between \emph{age}, \emph{body fat}, \emph{hip circumference} as well as \emph{menstrual period}.
 }
  \label{fig:Application}
\end{figure*}
%%
In this section, we describe how we applied the \emph{Regression Cube} to two epidemiological data sets.
%%
The hepatic steatosis data set was analyzed using data mining algorithms, yielding risk groups, which we now analyze further.
%%
Also, we try to reproduce the prior results from this analysis as proof-of-concept of our method.
%%
The breast fat data set is the foundation for an explorative analysis towards the influencing parameters of the parenchyma tissue of the female breast.
%%

@@

% OK so?
The third participant is \emph{TI}, a statistician with focus on epidemiology (8 years of experience) assesses the statistical reliability of the tool and the underlying methods without a focus on a specific data set.
%%
\subsection{The Hepatic Steatosis Data Set}
We employ the data set used in \cite{Niemann2014} to identify predictive features w.r.t. the reversible hepatic steatosis disorder.
%%
The dichotomous target feature is derived from the liver fat concentration measured using MRI scans.
%%
% ok?  to indicate absence or presence of the disease.
Liver fat concentrations of no more than 10\% are mapped to the `negative' class; values greater than 10\% are mapped to the `positive' class to indicate absence or presence of the disease.
%%
The data set contains labels for 578 participants for which the MRI data was available to that point.
%%
The MRI scans for each subject have been introduced in SHIP-2, so the target feature is only present for one of three study waves.

@@

In \cite{Niemann2014}, the authors show different class distributions of liver fat concentrations of women and men.
%%
For women, an association between age and liver fat was identified.
%Also, for women they identify that there is a association between age and liver fat concentration.
%%
An appropriate cut-off value of 52 years was set, yielding the most homogeneous class distribution within the resulting subsets.
%%
% ist das richtig? nicht female<52,male, female>52?
Based on these observations, we perform our analysis on three sub-populations: \emph{females}, \emph{males} and \emph{females older than 52 years}.
%%
\subsection{The Breast Fat Data Set}
The breast fat data set was compiled to find associations between the parenchyma tissue proportion in the female breast compared to other features in the data.
%%
The ratio between parenchyma and cellular connective tissue (breast density) has been shown to be associated with breast cancer.
%%
Studies describe a four to five times increased risk of getting breast cancer for participants with a breast density above 50\% \cite{Mccormack2006}.

The data comprises 1,186 female subjects (368 from \texttt{SHIP-2}, 818 from \texttt{SHIP-TREND-0} cohort).
%%
It contains 231 dimensions, holding information about somatometric features (e.g., body size or weight) consumption behavior (e.g., alcohol or tobacco), personal and medical history (e.g., occupation or prior diseases), women-specific features (e.g., number of born children or contraception type) as well as mammography features (e.g., fat content or parenchyma tissue proportion to volume).
% It contains 231 dimensions, holding information about
% \begin{itemize}
% 	\item \emph{somatometric features}, e.g., body size, weight or BMI,
% 	\item \emph{lifestyle features}, e.g., alcohol/tobacco consume,
% 	\item \emph{personal history}, e.g., occupation, marital status,
% 	\item \emph{medical history}, e.g., current or prior diseases or laboratory values, such as cholesterol levels,
% 	\item \emph{women specific features}, e.g., number of born children, contraception type or hormone replacement therapy and
% 	\item \emph{mammography features} derived from the image data.
% \end{itemize}
%%
The latter were derived from manually segmented MRI data for each subject carried out by radiologists.
%%
%%
%\com{Describe Parenchym tissue and its relation to breast cancer.}
%\com{Describe how the data was acquired and preprocessed}

@@

All features related to the mammography attributes were stored in an additional file.

@@

We refer to each analysis step with regard to its belonging in the VA-Mantra (recall Sec.~\ref{sec:Workflow}).

@@

Therefore, \emph{UN} started the [\textbf{1.AF}] step using the dichotomized MRI fat liver concentration and the formula $mrt\_liverfat\_s2 \sim X + Y + Z$ for \emph{male} subjects.

@@

The Zoom, Filter and Analyze Further Step [\textbf{3.ZF}] was realized by slicing through the cube using the mouse input to inspect the hotspots.

@@

Analyzing the mosaic plot [\textbf{4.DD}] revealed high correlations for somatometric features, hepatic steatosis indicator features as well as laboratory values, such as \emph{creatinine} (used renal retention parameter) and \emph{uric acid} (among others used as gout and diabetes risk factors) magnitudes.

@@

Due to the difference between our regression model approach and the decision tree approach presented in \cite{Niemann2014}, a complete matching set of correlating features is not expected.

@@

\textbf{Analysis of non-discretized target feature.}
%%
%Since our method can analyze numerical target features, as opposed to decision trees used by \emph{UN} before, the analysis was conducted again for the not-dichotomized target using the same formula.
Since our method can assess numerical target features, the analysis was conducted again for the non-dichotomized target using the same formula.
%%
The 3D cube representation already showed lower $R^2$ values in general. However, the analysis is now based on linear regression and the $R^2$ values cannot be directly compared.
%%
The correlation hotspots matched with the ones from the dichotomous target, but were generally lower ($R^2$ of $0.37$ for somatometric features as opposed to $0.58$).
%%
We assume that the bias introduced by dichotomizing the fat liver content enforces the findings of liver diseases, while using the numerical features is less expressive.

\textbf{Interleukin-6 correlation with liver fat.}
%%
During the analysis, one hotspot was always observable in the [\textbf{2.SI}] and [\textbf{3.ZF}] incorporating a high \emph{Interleukin-6} (\emph{IL-6}) correlation with liver fat values (Fig.~\ref{fig:Application} b).
%%
The correlation was high for both the dichotomized and continuous target feature.
%%
The literature described relations between \emph{IL-6} and liver cancer \cite{He2013} as well as chronic liver diseases \cite{Streetz2003}.
%%
For mice, strong effects of \emph{IL-6} with hepatic steatosis were described \cite{Hong2004}.
%%
The finding is subject of further analysis.
%%
\subsection{Case 2: Hypothesis-free Analysis of the Breast Cancer Data Set}
%%
The analysis aims to find relationships on the breast fat data using mammography analysis features.
%%
Relationships between the share parenchyma tissue on the overall breast volume are of high interest \cite{Mccormack2006}.
%%
The [\textbf{1.AF}] was started by \emph{KH} using the default formula for hypothesis-free analysis ($Z \sim X + Y$).
%%
At first, she was interested in correlations with the \emph{parenchyma tissue} percentage, which was selected through the drop-down for the z-axis [\textbf{2.SI}].
%%
She observed strong correlations with \emph{age}, \emph{body fat percentage}, \emph{hip} and \emph{waist circumference} as well as \emph{menstrual period} or \emph{pregnancy status} as expected (Fig.~\ref{fig:Application} right).
%%
Women with higher \emph{body fat} also have a larger \emph{breast fat percentage}, which also correlates with other somatometric features.
%%
\emph{Age} is a strong influencing factor, as breast tissue and subsequently the parenchyma tissue degrades over time.
%%
\emph{KH} proceeded using [\textbf{3.ZF}] and [\textbf{4.DD}] to check for relationships for different target features, such as current \emph{hormone replacement therapy}, \emph{BI-RADS} (classification of the mammography findings) as well as different diseases, such as \emph{diabetes} or \emph{gout}.
%%
She observed relationships matching her expectations and expert knowledge.
%%
One unexpected relationship was observed between \emph{breast lesions} and \emph{menstruation cycle} towards \emph{spiral contraception} ($R^2$ of $0.77$).
%%
\emph{KH} proceeded with a detailed analysis of the parenchyma tissue.
%\emph{KH} derived at the hypotheses: \emph{Somatometric features impose the major influences on the parenchyma tissue.}
%%

\textbf{Detailed Breast Parenchyma Analysis.}
%%
The analysis was conducted by calculating the cube $Parenchyma\_Percentage \sim X + Y + Z$ [\textbf{1.AF}].
%%
Using the 3D cube, \emph{KH} observed several hotspots [\textbf{2.SI}].
%%
Navigating to them using the slicing facility of the 3D visualization [\textbf{3.ZF}] highlighted features of high influence, such as image-derived features, as \emph{glandular tissue density} and \emph{parenchyma segmentation} metrics.
%%
Also, strong correlations were observed in the \emph{diabetes} slice, confirming expectations of \emph{KH} w.r.t. its strong influence to the parenchyma tissue.
%%
A surprising finding were the strong correlations with \emph{kidney disorder} ($R^2$ values around $0.9$).
%%
The [\textbf{4.DD}] analysis, however, showed only 8 subjects with \emph{kidney disorder}.
%%
Too few subjects impose the risk of a biased finding.
%%
The correlation was noted and will be further investigated in the near future using an extensive data set.
%%
Lastly, \emph{KH} assessed the influence of contraception-related features, such as use of \emph{birth control pills} or the \emph{spiral}, but found no significant correlations with the parenchyma tissue.
%%
Other consumption behavior features, such as \emph{alcohol intake} also yield no elevated $R^2$ values.
%%
\emph{KH} remarked that these features are believed to have an impact on the parenchyma tissue, but they are less reliable, since they are self-reported.
%%
%mamma anamnesis breast fed, contraception (spirale, pille), alcohol intake
%Lastly, \emph{KH} assessed the influence of image-derived features on the parenchyma tissue.
%%
%She observed expected results 
%Using the slicing mode to fast 
%Hypotheses
%Detailed analysis: Cholesterin, influence of image-derived feautres, classification of the BI-Rads (how well did the manual segmentation perform)
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
Our method was well received among the domain experts.
%%
For the first time, they were able to derive an overview visualization custom-tailored to underlying assumptions.
%%
\emph{KH} noted the ease of use, which "\emph{converts data sets into a feasible form}".
%%
She highlighted the efficiency of combining fast target feature selection with visually highlighting interesting results, enabling rapid analysis cycles.
%%
To get nearly similar results, she had to spend hours using \texttt{SPSS} and potentially missed interesting hotspots during this process.
%She noted that she usually would have to spend hours using \texttt{SPSS} to assess the data in the detail we presented.
%%
\emph{TI} highlighted the ability to simultaneously analyze thousands of regression models while maintaining little time expenses for rating them.
%The novel way of modeling knowledge to derive hypothesis driven overview visualizations allow for detailed analysis of target features, either disease indicators or even image-derived features.
%%
%TI commented on the

@@

\textbf{Overview Visualizations are Preferred over Black-box Methods.}
%%
Explorative analysis based on the data gains importance in epidemiology with increasing data set complexity.
%While the classical epidemiological workflow represents the foundation of the research in this area, explorative analysis based on the data gains importance.
Results from automatic `black-box' methods, such as data mining algorithms are more often obscure to the expert.
%%
Findings and hypotheses derived through overview visualizations, however, are met with more confidence, because the users actually observed the behavior themselves.
%%
The participation and steering of the analysis using human pattern detection and expert knowledge is preferred.
%%
%Incorporating data mining algorithms with overview visualizations of statistical measures allow for reasoning about risk factors using underlying epidemiological data sets rather than predefined hypotheses.
%%
%We  a higher confidence of an expert in a result he or she actually observed in the data rather 
%One major advantage we observed is the confidence of an expert in the results he or she actually observe.
%%
Observing \emph{expected} correlations matching the expert knowledge strengthens the confidence in the method and, subsequently, in the hypotheses generated from unanticipated relationships.
%Observing \emph{expected} correlations, which match the expert knowledge, strengthens the confidence in the method and, subsequently, in the hypothesis generated from unanticipated relationships.
%%
%\com{Include the human into the analysis provides more confidence into the result than data mining algorithms. People believe more in what they see, rather than in what they are presented by an algorithm. Regression cubes incorporate this by showing known associations as well as new ones. Preattentive feature display (either through fast slicing through a cube or by showing hot spots) helps.}

%%
\textbf{Using non-discretized features reduces information bias.}
%%
Discretization reduces the information space and introduces bias into the data and is therefore avoided in epidemiological research as possible.
%%
In contrast to many data mining algorithms, our method allows to use the concurrent analysis of heterogeneous data types.
%%
Investigations of the hepatic steatosis data set with both numerical and dichotomized liver fat values showed comparable results but slightly differed when it comes to details. %, but differences  %differing in detail.
%%
The overall explanatory power towards the numerical feature was lower, supporting the hypothesis that the dichotomized target feature already models knowledge to bias the results towards the expected result.
%%
%Our investigation regarding the hepatic steatosis data set confirms differences in results
%%
%\com{Comment on the lesson learned from the last VAST Paper - "Usage of different categorizations depending on expected outcome." Now we have continuos data.}
%%

\textbf{Attention steering is crucial.}
%%
Important events have to be highlighted in overview visualizations to steer the user's attention to interesting parts of the data.
%%
Poor guidance potentially leads to overlooked relationships.
%%
We found supporting mini-map visualizations, such as the 3D cube, most useful for this purpose, e.g., for highlighting differences rather than displaying absolute values (Fig.~\ref{fig:Application}).
%%
%Foe example, formulas with a static target feature (e.g. $Breast\_Cancer ~ X + Y + Z$) are rendered by highlighting differences rather than absolute values in the 3D cube visualizations (see Fig.~\ref{fig:Application} left) by comparing local values with the mean along the z-ray.
%%

@@

\section{Summary and Conclusion}
%%
We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Model expert knowledge using regression formulas allows for a \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D mosaic plot for each cube slice.
%%
Details on demand for each model allow for a detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.

@@

We published all associated codes and provide an analysis platform open to heterogenous data types.
%%
We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.

% \com{Rewrite, focus on precise definition on what was done and why it is awesome.}
% We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
% %%
% \emph{Regression Cubes} are based on eponymous regression models, which allow to model domain knowledge about the data (e.g. influencing features on a disease) using formulas.
% %%
% Alternatively, \emph{Regression Cubes} can be calculated using default formulas to assess the explanatory power towards each feature in the data set.
% %%
% These two approaches allow both for a \emph{hypothesis-free} and \emph{hypothesis-based} explorative analysis.
% %%
% Correlation-based feature selection reduces the amount of calculations using the target feature by focusing on the important regression models.
% %%
% The prototype described in this work was developed using state of the art web technologies, free services and open source libraries.
% %%
% All code associated to this project is also open source.
% %%
% Two case studies revealed that the presented approach enables to reproduce knowledge extracted using decision-tree-based data mining methods on a hepatic steatosis data set, as well as producing new hypotheses by deriving insight into influencing factors on breast fat tissue in a explorative analysis.
% %%
%
% As future work, we want to introduce more regression models to the data set, which model different correlation types.
% %%
% We also want to extend the cube to time-dependent data by expanding the difference-cube approach.
% %%
%
% The method was very well received with our project partners, allowing them for the first time to retrieve an overview visualization of features, which does not highlights their pairwise correlations, but rather their explanatory power towards a target feature.
% %%
% The observed relationship are now subject of detailed statistical analyses.
% %%
% Our provided system not restricted to the epidemiological domain, using the provided instance, everybody can upload their own data or even set up their own \emph{Regression Cube} server cluster.
% %%
% We believe in the power of opening up knowledge discovery to allow a heterogenous group of domain experts to derive insight into their data and support the notion of open science.
%%
%%\paragraph{Future Work.}
%%\com{Implementing of regression analysis on the graphics card.}
%%
\begin{small}
	%\acknowledgments{Omitted due to blind review}
\acknowledgments{SHIP is part of the Community Medicine Research net of the University of Greifswald, Germany, which is funded by the Federal Ministry of Education and Research (grant no. 03ZIK012), the Ministry of Cultural Affairs as well as the Social Ministry of the Federal State of Mecklenburg-West Pomerania. Whole-body MR imaging was supported by a joint grant from Siemens Healthcare, Erlangen, Germany and the Federal State of Mecklenburg-Vorpommern. The University of Greifswald is a member of the ‘Centre of Knowledge Interchange’ program of the Siemens AG. This work was supported by the DFG Priority Program 1335: Scalable Visual Analytics. This work was supported by the federal state of Saxony-Anhalt under grant number `I 60' within the Forschungscampus STIMULATE.}
\end{small}
\clearpage
\newpage
\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{bibliography}
\end{document}
